A Reproducible Data Analysis Workflow with RMarkdown, Git, Make & Docker

### 2020-05-12 manuscript.Rmd
added worcs references

While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow. There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed. The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr]. The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies. The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker. The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of Docker. `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required. `holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited. There are several alternatives to Docker that manage dependencies on R packages. [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies. Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint]. The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited.

>>>>

While our approach was designed to scale well with the complexity of a computing-intense project, we realise that this flexibility may not be straightforward to integrate into a researcher's everyday workflow. There are R packages that implement parts of our workflow and, thus, lower the threshold for adoption when the full flexibility provided by our workflow is not needed. The use of R Markdown within a project, tracked with Git can be simplified with the [`workflowR`](https://github.com/jdblischak/workflowr) package [@workflowr]. The [`drake`](https://github.com/ropensci/drake) package [@drake] is directly inspired by Make and takes an R centric approach, making it especially suited for projects only involving R, but can also handle external dependencies. The [liftR ](https://cran.r-project.org/web/packages/liftr/vignettes/liftr-intro.html) package [@liftr] and the [`holepunch`](https://karthik.github.io/holepunch/) package [@holepunch] automatize the use of Docker. The former is perfectly compatible with the described workflow and we recommend it to users  who are not comfortable with command-line use of Docker. `holepunch` uses [binder](https://mybinder.org) [@binder2018] to move the analysis to the cloud, so that no local installation of Docker is required. `holepunch` is well suited for simple analysis with low computational demands, because binder's  memory and computing time is limited. There are several alternatives to Docker that manage dependencies on R packages. [`renv`](https://rstudio.github.io/renv/articles/renv.html) [@renv] is a way to freeze package version via local copies of packages in the project, but it does not gurantee a given base R version or system dependencies. Similar approaches are taken by [`jetpack`](https://github.com/ankane/jetpack) [@jetpack], [`miniCRAN`](https://cran.r-project.org/web/packages/miniCRAN/vignettes/miniCRAN-introduction.html) [@miniCRAN] and [`checkpoint`](https://cran.r-project.org/web/packages/checkpoint/vignettes/checkpoint.html) [@checkpoint]. The package [`reprex`](https://github.com/tidyverse/reprex) [**repr**oducible **ex**ample, @reprex] is also worth noting, but its scope is limited. A particular noteworthy approach is provided by the [`worcs`](https://cran.r-project.org/web/packages/worcs) package [@worcs, @worcspaper], which is an R project template that creates a standardized file structure for code and data supporting version management with git, package management with renv and dynamic document generation with R Markdown. We acknowledge that worcs is much easier to install and provides a one-click solution for the creation of reproducible projects. It achieves a high standard of reproducibility but does not guarantuee full computational reproducibility and is limited to dependency management within the R environment.
